---
title: "ACS updates"
description: New or changed ACS tables by year
vignette: >
  %\VignetteIndexEntry{ACS updates}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    comment: '#>'
    collapse: true
execute:
  warning: false
  message: false
  echo: false
fig-dpi: 150
fig-width: 6
---


```{r}
#| label: libs
#| message: false
library(cwi)
library(dplyr)
library(purrr)
```

```{r}
#| label: compare-yrs
compare_yrs <- function(yr, ret = FALSE) {
    yrs <- rlang::set_names(c(yr - 1, yr))
    short_yrs <- rlang::set_names(as.character(yrs), sprintf("_%s$", letters[seq_along(yrs)]))
    tbls <- purrr::map(yrs, cwi:::clean_acs_vars, "acs5")
    tbls <- purrr::map(tbls, mutate, table_num = stringr::str_remove(name, "_.+$"))
    tbls <- purrr::map(tbls, distinct, concept, table_num)
    tbls <- purrr::map(tbls, mutate, concept = stringr::str_squish(tolower(stringr::str_replace_all(concept, "20\\d{2}", "20xx"))))

    purrr::map(tbls, ~ paste("Number of rows:", nrow(.)))

    out <- purrr::reduce(tbls, dplyr::full_join, by = "table_num", suffix = sprintf("_%s", letters[1:2]))
    out <- dplyr::filter(out, concept_a != concept_b | is.na(concept_a) | is.na(concept_b))
    out <- dplyr::rename_with(out, \(x) stringr::str_replace_all(x, short_yrs), .cols = dplyr::matches("concept_[a-z]$"))
    # fix misspelling in 2022
    out <- dplyr::mutate(out, dplyr::across(dplyr::matches("^concept"), \(x) stringr::str_replace(x, "major the population", "major for the population")))

    print(knitr::kable(out))

    if (ret) {
        return(out)
    }
}
```

# 2023 ACS

```{r}
#| label: c2023
#| results: asis
compare_yrs(2023)
```

# 2022 ACS

```{r}
#| label: c2022
#| results: asis
compare_yrs(2022)
```



# 2021 ACS

```{r}
#| label: c2021
#| results: asis
compare_yrs(2021)
```


# 2020 ACS

```{r}
#| label: c2020
#| results: asis
compare_yrs(2020)
```


Not a lot new in the 2020 ACS.

----

# 2019 ACS

## New tables

The 2019 ACS included [reengineering of a few tables](https://www.census.gov/programs-surveys/acs/technical-documentation/table-and-geography-changes/2019/5-year.html), and a few new additions. Some are clarifications of phrasing to both questions and tables, such as time leaving home for work. More substantial are the changes in how family structure has been phrased to be more inclusive of same-sex and unmarried couples--see the most recent [working paper](https://www.census.gov/content/census/en/library/working-papers/2017/demo/SEHSD-WP2017-28.html) on this.

This is just a quick check of what's new or changed.

```{r}
#| label: c2019
#| results: asis
change19 <- compare_yrs(2019, ret = TRUE)
```

## Variables in the tables related to gender & family structure:

```{r}
#| label: c2019-gender
#| results: asis
acs19 <- cwi:::clean_acs_vars(2019, "acs5") |>
    mutate(table_num = stringr::str_remove(name, "_.+$"))
change19 |>
    filter(grepl("(HOUSEHOLD|SPOUSE)", concept2019)) |>
    inner_join(acs19, by = "table_num") |>
    split(~table_num) |>
    purrr::map(function(x) split(x, x$concept)) |>
    purrr::map_depth(2, separate_acs) |>
    purrr::map_depth(2, select, variable = name, matches("^x\\d+")) |>
    purrr::map_depth(2, knitr::kable)
```





