---
title: "Basic workflow"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	collapse = TRUE,
	comment = "#>",
	fig.width = 7,
	fig.height = 4
)
```

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(camiller)
library(cwi)
```

The tables I'm working with are `B01003`, total population; `B03002`, race and Latino ethnicity; and `B25003`, housing tenure. It's easiest to save these in a named list, then map over the list calling `multi_geo_acs` for each table number.

```{r}
year <- 2016
table_nums <- list(
  total_pop = "B01003", 
  race = "B03002",
  tenure = "B25003"
)
```

I'm pulling out the entries in the `cwi` dataset `regions` (a list) to only include the Greater New Haven-area ones. Then I fetch the ACS tables, and filter them to only have geographic levels other than town, or to only be towns in Greater New Haven.

```{r}
gnh_regions <- regions[c("Greater New Haven", "New Haven Inner Ring", "New Haven Outer Ring")]

gnh_data <- table_nums %>%
  map(~multi_geo_acs(table = ., 
                     year = year, 
                     towns = "all", 
                     regions = gnh_regions, 
                     counties = "New Haven", 
                     state = "09")) %>%
  map(~filter(., NAME %in% gnh_regions$`Greater New Haven` | level != "4_towns"))
```


```{r}
gnh_data$total_pop
```

The total population data is very straightforward, as it only has one variable, `B01003_001`. The tibble returned has the GEOID, except for custom geographies like regions; the name of each geography, including the names of each region; the variable codes; estimates; margins of error at the default 90\% confidence level; the geographic level, numbered in order of decreasing size; and the counties of the towns.

The race and ethnicity table will require some calculations, using the brilliantly-titled `camiller` package. Join the `race` tibble with the `acs_vars` dataset to get variable labels. Oftentimes, these labels need to be separated by their `"!!"` delimeter. Group by the geographic level, county, and name, then call `camiller::add_grps` with a list of racial groups and their labels' positions in the `label` column. This gives estimates and, optionally, margins of error for aggregates; `camiller::calc_shares` then gives shares of each group's estimate over the `"total"` denominator.

```{r}
gnh_data$race %>%
  label_acs() %>% 
  group_by(level, county, NAME) %>%
  add_grps(list(total = 1, white = 3, black = 4, latino = 12, other = 5:9), group = label) %>%
  calc_shares(group = label, denom = "total")
```

With the tenure table, it's easiest to separate the labels by `"!!"`. Here the table can be wrangled into shares of households that are owner-occupied.

```{r}
homeownership <- gnh_data$tenure %>%
  label_acs() %>%
  separate(label, into = c("total", "tenure"), sep = "!!", fill = "left") %>%
  select(level, name = NAME, tenure, estimate) %>%
  filter(tenure != "Renter occupied") %>%
  group_by(level, name) %>%
  calc_shares(group = tenure) %>%
  filter(!is.na(share))

homeownership
```

`geo_level_plot` gives a quick visual overview of the homeownership rates, highlighting town-level values.

```{r}
homeownership %>% 
  geo_level_plot(value = share, hilite = "darkslateblue", type = "point")
```

`acs_quick_map` gives a quick map sketch of the rates.

```{r}
homeownership %>% 
  acs_quick_map(value = share, n = 6, color = "black", size = 0.4, title = "Homeownership, Greater New Haven")
```

