---
title: "crosstabs"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: >
  %\VignetteIndexEntry{crosstabs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	fig.height = 4,
	fig.width = 7,
	message = TRUE,
	warning = FALSE,
	collapse = TRUE,
	comment = "#>",
	dpi = 150
)
```

```{r libs, message=FALSE, warning=FALSE}
library(dplyr)
library(stringr)
library(cwi)
```

This is a quick look at using the crosstab-related functions being finished up for what will likely be `cwi` v.0.2.0. There are 3 test files of crosstabs that ship with this package for internal testing and illustration. This vignette recreates the data in the `cws_demo` dataset. 

`read_xtabs` reads a crosstab file, but keeps it ugly—you can then transform it into a ready-for-analysis data frame with `xtab2df`. Below all the survey data, the file has a table of weights. It's hard to know where that table might be and how to separate it from the data tables. `read_xtabs` reads until a marker that distinguishes between the data and the weights; `read_weights` reads everything below that marker.

The data for the first question looks like this:

```{r read_xtabs}
xtab_in <- read_xtabs(system.file("extdata/test_xtab2015.xlsx", package = "cwi"), year = 2015)

xtab_in[1:7, ]
```

Reshape the crosstabs into something you'd actually want to work with using `xtab2df`. That snippet from the first chunk now looks like this (for the sake of space, just the overall values and values by income):

```{r xtab2df}
cws <- xtab2df(xtab_in)

cws %>%
  filter(code == "Q8", category %in% c("Greater New Haven", "Income"))
```

Say we want to collapse those income brackets into just 3 groups. For that, use the weights table...and fix the group labels in the weights table to match the ones from the data (I'm only doing this for income, but education often needs reconciling as well). Obviously there won't be weights for the total group; in doing weighted means, the easiest fix is to replace NAs with 1. (FYI, groups Yes and No correspond to the category of having children in the household.)

```{r read_weights}
wts_in <- read_weights(system.file("extdata/test_xtab2015.xlsx", package = "cwi"))
wts_in

wts <- wts_in %>%
  mutate(group = group %>%
           str_replace("(?<=\\d) to (?=[\\$\\d])", "-") %>%
           str_replace_all(",000", "K") %>%
           str_replace("Less than (?=\\$)", "<") %>%
           str_replace(" or more", "+")) 

income_lvls <- list(
  "<$30K" = c("<$15K", "$15K-$30K"),
  "$30K-$100K" = c("$30K-$50K", "$50K-$75K", "$75K-$100K"),
  "$100K+" = c("$100K-$200K", "$200K+")
)

cws %>%
  filter(code == "Q8", category %in% c("Greater New Haven", "Income")) %>%
  left_join(wts, by = "group") %>%
  mutate(across(category:response, forcats::as_factor)) %>%
  tidyr::replace_na(list(weight = 1)) %>%
  collapse_n_wt(c(code, category:response), .lvls = income_lvls)
```

We generally want to remove non-answers (don't know, refused) from the denominator, then rescale the response values. Usually those non-answers only account for 1 or 2 percentage points and don't change values much, but occasionally they're large enough to make a noticeable difference. For consistency, I always remove non-answers, even when they're small.

With non-answers:

```{r nons1}
cws %>%
  filter(code == "Q1", category %in% c("Greater New Haven", "Gender"))
```

Without (using `sub_nonanswers` defaults):

```{r nons2}
cws %>%
  filter(code == "Q1", category %in% c("Greater New Haven", "Gender")) %>%
  sub_nonanswers()
```

