---
title: "Shapefiles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Shapefiles}
  %\usepackage[utf8]{inputenc}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 3,
  dpi = 150
)
```

```{r libraries, message=FALSE}
library(dplyr)
library(ggplot2)
library(cwi)
if (require(sf)) {
  sf::sf_use_s2(FALSE)
}
```

`cwi` ships with several shapefiles in [`sf`](https://r-spatial.github.io/sf/) (simple features) format. These include:

* Shapes straight from [TIGER](https://www.census.gov/geo/maps-data/data/tiger.html) via [`tigris`](https://github.com/walkerke/tigris):
    + `town_sf`
    + `tract_sf`
* Shapes put together at DataHaven based on official city neighborhood boundaries:
    + `bridgeport_sf`
    + `hartford_sf`
    + `new_haven_sf`
    + `stamford_sf`

# `sf`

`sf` objects are easy to plot either out of the box with `sf::plot.sf()`, or with `ggplot2::ggplot()` as of `ggplot2` version 3.0.0. Use with `ggplot()` means a `geom_sf()` can be overlayed on a basemap from `ggmap`, as is done in the neighborhood examples.

`sf` is well-documented, and uses `dplyr`-type functions to allow for mutating, filtering, summarizing, joining, and most other data wrangling functions. See the `sf` vignettes for help getting started: `vignette(package = "sf")`.

`sf` data frames look and act like other data frames, but with shape information attached.


# Statewide

A basic plot of `cwi::town_sf`, with some `ggplot` aesthetics set, graticule lines turned off with `coord_sf(ndiscr = F)`, and a blank theme.

```{r towns, out.width="100%"}
ggplot(town_sf) +
  geom_sf(fill = "gray70", color = "white", size = 0.5) +
  coord_sf(ndiscr = FALSE) +
  theme_void() +
  ggtitle("Connecticut towns")
```

With some `dplyr` operations. For example, `group_by()` followed by `summarise()` conducts a union on each group's geometries, using `sf::st_union()` in its back-end.

```{r town_regional, out.width="100%"}
town_sf %>%
  filter(name %in% regions[["Greater New Haven"]]) %>%
  group_by(area = ifelse(name == "New Haven", "New Haven", "New Haven suburbs")) %>%
  summarise() %>%
  ggplot() +
  geom_sf(aes(fill = area), color = "white") +
  coord_sf(ndiscr = FALSE) +
  scale_fill_manual(values = c("New Haven" = "mediumorchid", "New Haven suburbs" = "gray70")) +
  theme_void() +
  theme(legend.position = "bottom") +
  ggtitle("Filtering and unioning shapes")
```

`cwi::tract_sf`, a fairly large shape containing all tracts in the state, as of 2020 redistricting.

```{r tracts, out.width="100%"}
ggplot(tract_sf) +
  geom_sf(fill = "gray70", color = "white", size = 0.1) +
  coord_sf(ndiscr = FALSE) +
  theme_void() +
  ggtitle("Connecticut tracts")
```

# Neighborhoods

Neighborhood shapes were made at DataHaven by splitting or grouping census tracts into neighborhood areas as defined by each city's government. In some cases, boundaries were slightly adjusted to fit tracts. Small neighborhoods may be grouped together for analysis. See documentation of these at `cwi::neighborhood_shapes` and crosswalk of tracts to neighborhoods at `cwi::neighborhood_tracts`.

```{r city_maps, message=FALSE, warning=FALSE, out.width="100%"}
list(
  "New Haven" = new_haven_sf, Stamford = stamford_sf,
  "Hartford/West Hartford" = hartford_sf, Bridgeport = bridgeport_sf
) %>%
  purrr::iwalk(function(shp, title) {
    p <- ggplot(shp) +
      geom_sf(color = "mediumorchid3", fill = "orchid3", alpha = 0.1) +
      geom_sf_text(aes(label = name), size = 2.5) +
      coord_sf(ndiscr = FALSE) +
      theme_void()
    print(p)
  })
```





